---
title: "covid 19"
author: "Tracy Alset"
date: "07/21/22"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:

```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

Based on the table, it looks like more females died than males, but that doesn't tell if dying is independent from sex. In order to answer this question we have to follow the proceures ahead and get some data that can help us get to the answer.


## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this


```{r}
kRdata1 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(kRdata1,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(kRdata1,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

Based on the frequency graph, it looks like more males died than females. Looking at the percentage graph, the percentage of males who died is almost double the percentage of females who died. So far we can't tell if the sex is affecting our data, so we keep analyzing using numerical data.





  
## Numerical results


## Inferential results 




## Conclusions

